---
title: "Untitled"
author: "Alice Anonymous"
date: "6/30/2020"
output:
  pdf_document:
    df_print: kable
  md_document:
    variant: markdown_github
    df_print: kable
knit: (function(input,encoding) {
        rmarkdown::render(
          input=input, 
          encoding=encoding, 
          output_dir="../",
          output_file='README.md',
          output_format='md_document');
        rmarkdown::render(
          input=input, 
          encoding=encoding,
          output_dir='../',
          output_file='Analysis.pdf', 
          output_format='pdf_document')})
---

```{r setup, include=F, echo=F}
library(knitr)
knitr::knit
```

# Abstract


A patient was considered positive for DMII or MDD if any of the following ICD-9 codes were recorded for that patient across all hospital admissions.

**_ICD-9 CM codes defining DMII_**

Non-specific code [250](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/default.htm) Diabetes mellitus

- [250.00](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.00.htm) DM type II or unspecified type w/o complication, not stated as uncontrolled)
- [250.02](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.02.htm) DM type II or unspecified type w/o complication, uncontrolled)
- [250.10](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.10.htm) DM type II or unspecified type w/ ketoacidosis, not stated as uncontrolled)
- [250.12](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.12.htm) DM type II or unspecified type w/ ketoacidosis, uncontrolled)
- [250.20](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.20.htm) DM type II or unspecified type w/ hyperosmolarity, not stated as uncontrolled)
- [250.22](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.22.htm) DM type II or unspecified type w/ hyperosmolarity, uncontrolled)
- [250.30](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.30.htm) DM type II or unspecified type w/ other coma, not stated as uncontrolled)
- [250.32](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.32.htm) DM type II or unspecified type w/ other coma, uncontrolled)
- [250.40](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.40.htm) DM type II or unspecified type w/ renal manifestations, not stated as uncontrolled)
- [250.42](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.42.htm) DM type II or unspecified type w/ renal manifestations, uncontrolled)
- [250.50](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.50.htm) DM type II or unspecified type w/ ophthalmic manifestations, not stated as uncontrolled)
- [250.52](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.52.htm) DM type II or unspecified type w/ ophthalmic manifestations, uncontrolled)
- [250.60](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.60.htm) DM type II or unspecified type w/ neurological manifestations, not stated as uncontrolled)
- [250.62](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.62.htm) DM type II or unspecified type w/ neurological manifestations, uncontrolled)
- [250.70](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.70.htm) DM type II or unspecified type w/ peripheral circulatory disorders, not stated as uncontrolled)
- [250.72](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.72.htm) DM type II or unspecified type w/ peripheral circulatory disorders, uncontrolled)
- [250.80](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.80.htm) DM type II or unspecified type w/ other specified manifestations, not stated as uncontrolled)
- [250.82](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.82.htm) DM type II or unspecified type w/ other specified manifestations, uncontrolled)
- [250.90](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.90.htm) DM type II or unspecified type w/ unspecified complication, not stated as uncontrolled)
- [250.92](http://www.icd9data.com/2014/Volume1/240-279/249-259/250/250.92.htm) DM type II or unspecified type w/ unspecified complication, uncontrolled)

**_ICD-9 CM codes defining MDD_**

Non-specific code [296](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/default.htm) Episodic mood disorders

  - [296.20](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.20.htm) Major depressive affective disorder, single episode, unspecified
  - [296.21](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.21.htm) Major depressive affective disorder, single episode, mild
  - [296.22](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.22.htm) Major depressive affective disorder, single episode, moderate
  - [296.23](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.23.htm) Major depressive affective disorder, single episode, severe
  - [296.24](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.24.htm) Major depressive affective disorder, single episode, severe
  - [296.25](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.25.htm) Major depressive affective disorder, single episode, in partial or unspecified remission
  - [296.26](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.26.htm) Major depressive affective disorder, single episode, in full remission      

Non-specific code [296.3](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.3.htm) Major depressive disorder recurrent episode

  - [296.30](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.30.htm) Major depressive affective disorder, recurrent episode, unspecified
  - [296.31](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.31.htm) Major depressive affective disorder, recurrent episode, mild
  - [296.32](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.32.htm) Major depressive affective disorder, recurrent episode, moderate
  - [296.33](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.33.htm) Major depressive affective disorder, recurrent episode, severe, w/o mention of psychotic behavior
  - [296.34](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.34.htm) Major depressive affective disorder, recurrent episode, severe, specified as with psychotic behavior
  - [296.35](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.35.htm) Major depressive affective disorder, recurrent episode, in partial or unspecified remission
  - [296.36](http://www.icd9data.com/2014/Volume1/290-319/295-299/296/296.36.htm) Major depressive affective disorder, recurrent episode, in full remission

# Load the dataset

The [data/study-population.csv](data/study-population.csv) file contains the results of a [query](sql/study-population.sql) on the MIMIC-III database that extracts patient information along with a one-hot encoding of ICD-9 CM diagnoses for Diabetes Mellitus Type II (DMII) and Major Depressive Disorder (MDD). 

```{r data, echo=T, message=F, warning=F}
library(tidyr)
library(dplyr)

set.seed(123)

POP_CSV <- '../data-raw/study_population.csv'

pop <- population <- read.csv(POP_CSV) %>%
  replace_na(list(dmii = 0, mdd = 0)) %>%
  mutate_all(na_if, '')

pos <- positive <- list(
  mdd = filter(pop, dmii == 0, mdd == 1), 
  dmii = filter(pop, dmii == 1, mdd == 0), 
  both = filter(pop, dmii == 1, mdd == 1) 
)

neg <- negative <- lapply(pos, function(sub) {
  pop %>% 
    filter(dmii == 0, mdd == 0) %>%
    sample_n(size=nrow(sub))
}) %>% setNames(names(pos))
```

```{r tbl1_tbl2, echo=F}
tbl1.cap <- sprintf('A sample of the study population (N = %s).', nrow(pop))

kable(
  data.frame(
    rbind(
      pos$mdd[1:2,],
      pos$mmii[1:2,],
      pos$both[1:2,],
      neg$both[1:2,]
    )
  ), 
  align='ccllccccc', caption=tbl1.cap
)
```

```{r tbl2, echo=F}
tbl2.cap <- 'Subsample size by diagnostic subcategory (DMII, MDD, or both).'
tbl2.col <- c('DMII(+)/MDD(+)', 'DMII(+)/MDD(-)', 'DMII(-)/MDD(+)')

kable(
  data.frame(
    nrow(pos$both), 
    nrow(pos$dmii), 
    nrow(pos$mdd), 
    row.names = c('Size (N)')
  ),
  align = 'c', col.names = tbl2.col, caption = tbl2.cap
)
```

# Chi-squared test
```{r xsq, echo=T, message=F, warning=F}

(ctab <- data.frame(
  dmii = c(
    sum(pos$dmii$gender == 'M'),
    sum(pos$dmii$gender == 'F')
  ),
  mdd = c(
    sum(pos$mdd$gender == 'M'),
    sum(pos$mdd$gender == 'F')
  ),
  both = c(
    sum(pos$both$gender == 'M'),
    sum(pos$both$gender == 'F')
  ),
  row.names = c('M', 'F')
))

chisq.test(
  ctab
)

install.packages('gplots')
library("gplots")
# 1. convert the data as a table
dt <- as.table(as.matrix(ctab))
# 2. Graph
balloonplot(t(dt), main ="housetasks", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)

install.packages('graphics')

mosaicplot(dt, shade = TRUE, las=5,
           type='deviance')

```
